Spatial distribution of the COVID-19 lockdown crime drop in England and Wales

### 2021-03-20 draft_paper.Rmd
anony process

---
title: "Spatial patterning of the COVID-19 lockdown crime drop in England and Wales"
author:
  - Samuel Langton
  - Anthony Dixon
  - Graham Farrell
csl: apa.csl
header-includes: \usepackage{float}
output:
  pdf_document: default
  word_document:
    reference_docx: reference_doc.docx
bibliography: refs.bib
---

```{r setup, include=FALSE, echo = F}
knitr::opts_chunk$set(echo = FALSE, eval = T) 
library(readr);library(kableExtra)
```

@@

Recent months have seen a wave of interest in describing and explaining how crime has changed in response to restrictions on mobility brought about to limit the spread of COVID-19. Studies have been undertaken in the United States [@campedelli2020exploring; @mohler2020impact; @piquero2020staying; @ashby2020initial], Australia [@payne2020covid; @payne2020property], the United Kingdom [@langton2021cs; @halford2020crime; @buil2020cybercrime] Canada [@hodgkinson2020show], Sweden [gerell2020minor], China (@borrion2020measuring) and Mexico [@estevez2020crime]. Generally speaking, findings have been consistent with opportunity perspectives and the mobility theory of crime during the pandemic [@langton2021cs; @halford2020crime]. That is, legally-enforced restrictions on daily activities, mobility and social interactions limited opportunities for criminality which in turn resulted in a macro-level decline in crime. As restrictions were relaxed, these opportunities reemerged, mobility patterns converged back to normality, and crime began to 'bounce back' -- often back to levels that would have been expected without the global pandemic and ensuing lockdowns.

@@

In England and Wales, early descriptive evidence from the first three months of lockdown suggests that previously high-crime areas may have experienced the steepest proportional declines [@dixon2020crime]. Increases in fly-tipping have also been linked to a small number of councils, with trends varying considerably between regions [@dixon2020fly]. However, there has not been a comprehensive decomposition of local variation and longitudinal (in)stability underpinning the nationwide lockdown crime drop and subsequent resurgence, or indeed an exploration of the opportunity structures characterizing the local areas which have driven macro-level changes.

@@

To examine localized instability in the lockdown crime drop, we make use of three data sources, namely, open police-recorded crime data, geographic boundaries from Ordnance Survey and the Office for National Statistics (ONS), and data sourced from the Open Street Map API. Each of these are now outlined in turn, followed by an outline of the methods deployed. Code to replicate the data downloads, handling, analyses and visualization reported here are openly available (anonymized GitHub repo link here).

@@

Without exception, the city centers and commercial districts of each city are characterised by 'high crime, major drop' clusters E and F, which in turn often neighbor the 'mid crime, mid drop' cluster D. By contrast, the 'low crime and stable' clusters tend to sit outside of the city centres: cluster B in the suburbs and the sparsely populated cluster A on the periphery of each urban conurbation. As such, changes in crime during lockdown appear to have a distinct geographic pattern: the nationwide drop was not experienced uniformly across space, but rather, particular areas including city centers appear to have driven the macro-level trend, with outer suburbs remaining fairly stable. For a detailed and interactive investigation of the spatial clusters identified for the whole of England and Wales, we refer readers to an openly available online map (anonymized link here).

@@

Here, we would note some weaknesses in the study. First, the crime measure of 'notifiable offences (excluding drugs)' aggregates away some detail in longitudinal trends. The crime type profile findings certainly go some way in addressing this, and we have now set a baseline from which further investigations can be conducted. However, further insight could be gained from replicating this analyses for major crime types which we know to have distinct opportunity structures, such as burglary and theft. Here, we would note that the generalized Gini coefficients by crime type (see Appendix) certainly suggest that there may be differences between crime types, as does preliminary findings using decile clusters

Secondly, the Open Street Map API used to quantify the prevalence of crime opportunities in each cluster is imperfect. Studies have found an association between contributions to Open Street Map and contextual characteristics such as poverty and distance from the city center -- factors that are also known to be associated with crime. Due to computational issues and API query limits, some features were also unobtainable on a national scale, such as the residential buildings footprints. While we would argue that Open Street Map is unique in its ability to offer open data on opportunity structures nationwide, we would encourage future research to verify these findings using other data sources.

>>>>

---
title: "Spatial distribution of the COVID-19 lockdown crime drop in England and Wales"
author:
  - Samuel Langton
  - Anthony Dixon
  - Graham Farrell
csl: apa.csl
header-includes: \usepackage{float}
output:
  word_document:
    reference_docx: reference_doc.docx
  pdf_document: default
bibliography: refs.bib
---

```{r setup, include=FALSE, echo = F}
knitr::opts_chunk$set(echo = FALSE, eval = F) 
library(readr);library(kableExtra)
```

@@

Recent months have seen a wave of interest in describing and explaining how crime has changed in response to restrictions on mobility brought about to limit the spread of COVID-19. Studies have been undertaken in the United States [@campedelli2020exploring; @mohler2020impact; @piquero2020staying; @ashby2020initial], Australia [@payne2020covid; @payne2020property], the United Kingdom [@langton2021cs; @halford2020crime; @buil2020cybercrime] Canada [@hodgkinson2020show], Sweden [gerell2020minor], China (@borrion2020measuring) and Mexico [@estevez2020crime]. Generally speaking, findings have been consistent with opportunity perspectives and the mobility theory of crime during the pandemic [@halford2020crime]. That is, legally-enforced restrictions on daily activities, mobility and social interactions limited opportunities for criminality which in turn resulted in a macro-level decline in crime. As restrictions were relaxed, these opportunities reemerged, mobility patterns converged back to normality, and crime began to 'bounce back' -- often back to levels that would have been expected without the global pandemic and ensuing lockdowns [@langton2021cs]

@@

In England and Wales, early descriptive evidence from the first three months of lockdown suggests that previously high-crime areas may have experienced the steepest relative declines compared to previous years [@dixon2020crime]. Increases in fly-tipping have also been linked to a small number of councils, with trends varying considerably between regions [@dixon2020fly]. However, there has not been a comprehensive decomposition of local variation and longitudinal (in)stability underpinning the nationwide lockdown crime drop and subsequent resurgence, or indeed an exploration of the opportunity structures characterizing the local areas which have driven macro-level changes.

@@

To examine localized instability in the lockdown crime drop, we make use of three data sources, namely, open police-recorded crime data, geographic boundaries from Ordnance Survey and the Office for National Statistics (ONS), and data sourced from the Open Street Map API. Each of these are now outlined in turn, followed by an outline of the methods deployed. Code to replicate the data downloads, handling, analyses and visualization reported here are openly available (https://anonymous.4open.science/r/46199882-cc5a-449d-9946-300b3132ab2e/).

@@

Without exception, the city centers and commercial districts of each city are characterised by 'high crime, major drop' clusters E and F, which in turn often neighbor the 'mid crime, mid drop' cluster D. By contrast, the 'low crime and stable' clusters tend to sit outside of the city centres: cluster B in the suburbs and the sparsely populated cluster A on the periphery of each urban conurbation. As such, changes in crime during lockdown appear to have a distinct geographic pattern: the nationwide drop was not experienced uniformly across space, but rather, particular areas including city centers appear to have driven the macro-level trend, with outer suburbs remaining fairly stable. For a detailed and interactive investigation of the spatial clusters identified for the whole of England and Wales, we refer readers to an openly available online map (anonymous for review).

@@

Here, we would note some weaknesses in the study. First, the crime measure of 'notifiable offences (excluding drugs)' aggregates away some detail in longitudinal trends. The crime type profile findings certainly go some way in addressing this, and we have now set a baseline from which further investigations can be conducted. However, further insight could be gained from replicating this analyses for major crime types which we know to have distinct opportunity structures, such as burglary and theft. Here, we would note that the generalized Gini coefficients by crime type (see Appendix) certainly suggest that there may be differences between crime types, as have preliminary findings using decile clusters [@dixon2020crime]. Secondly, the Open Street Map API used to quantify the prevalence of crime opportunities in each cluster is imperfect. Studies have found an association between contributions to Open Street Map and contextual characteristics such as poverty and distance from the city center -- factors that are also known to be associated with crime. Due to computational issues and API query limits, some features were also unobtainable on a national scale, such as the residential buildings footprints. While we would argue that Open Street Map is unique in its ability to offer open data on opportunity structures nationwide, we would encourage future research to verify these findings using other data sources.
